---
output: github_document
---
# Purpose

This README captures all my code which I had used in the creation of my thesis for 2025.

```{r Setup, include=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
pacman::p_load(tidyverse, foreign, haven, estimatr)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

# Running initial OLS on data

-   An OLS model was created to test if the variables are somewhat significant in the first place.
-   It helps to inform the general direction of the project and to establish any relationships from early on.

## Pull Data into R

```{r Pulling data into R}
# pull W1 data from the adult questionnaires
download.file(
    'https://www.dropbox.com/scl/fi/3s45c3flubpx0ngbz4qjq/Adult_W1_Anon_V7.0.0.dta?rlkey=oa17mxoimb3gqdo7x7jh90elu&st=dpzn40ah&dl=1',
    './data/w1_adult_anon.dta',
    mode = "wb"
)

w1 <- read_dta('./data/w1_adult_anon.dta')

# pull W4 data from adult questionnaires
download.file(
    'https://www.dropbox.com/scl/fi/3ggfkjhj55z09i11e40ft/Adult_W4_Anon_V2.0.0.dta?rlkey=wjrlq33ggrohe5o8z3mpjsrda&st=ul2l4ay9&dl=1',
    './data/w4_adult_anon.dta'
)
w4 <- read_dta('./data/w4_adult_anon.dta')
```

## Clean and wrangle data

-   Obtain W1 data and clean to include only variables which proxy for conscientiousness and other relevant variables

```{r Cleaning W1}
clean_w1 <- w1 %>% 
    select(pid, w1_a_dob_m, w1_a_dob_y, w1_a_gen, w1_a_unemmn, w1_a_hllfexer, w1_a_emohope, w1_a_aspen, w1_a_edter, w1_a_hllfsmk, w1_a_edintter, w1_a_wbsat) %>% # selecting our questions which proxied for measures of conscientiousness
    mutate(Age = 2008 - w1_a_dob_y) %>% # creating age variable 
    filter(Age >= 14 & Age <=23) %>% # filter for those aged 14-23 at time of interview
    select(-c(w1_a_dob_m, w1_a_dob_y)) # remove unnecessary data
```

-   Now want to clean W4 data to obtain W4 wages, employment, and schooling

```{r Cleaning W4}
clean_w4 <- w4 %>% 
    select(pid, w4_a_gen, w4_a_dob_y, w4_a_em1, w4_a_em1pay, w4_a_edschgrd) # collecting columns with info on employment status, wages and schooling level
```

Tips:

-   Include common sense test
-   Include dummy variable for missing values of conscientiousness, start with 4/5 most promising then add more if needed.
-   Use PCA/MCA for constructing measure. Do some research on this

## Joining W1 and W4

```{r Joining W1 and W4}
joined_w1_w4 <- clean_w1 %>% 
    left_join(clean_w4, by = "pid")
```

## Add a common sense test

This will help catch if there are errors in the data, e.g. does reported gender change from W1 to W4.

```{r Common sense test}
test1 <- joined_w1_w4 %>% 
    filter(!is.na(w1_a_gen) & !is.na(w4_a_gen)) %>% # removes entries where gender was missing in either w1 or in w4
    mutate(match = w1_a_gen == w4_a_gen) # returns TRUE/FALSE from logical test to see if the genders in W1 and W4 match for each observation

unique(test1$match) # checking if we have only TRUE values
```

Test shows that data is not flawed!! :)

(WITH OLD GENDER VARIABLE) 3 077 entries were successfully joined and had passed the common sense test. This is different from the `joined_w1_w4` sample size (4 148 observations) due to the exclusion of those entries with missing variables. 1 071 variables drop out because of a missing gender variable in either W1 or W4. Therefore must try with the `best_gen` variable as this provides a better measure for test.

## Trying common sense test with best gender

```{r Pull in different datasets}
download.file(
    'https://www.dropbox.com/scl/fi/etluqr514bxyzigroy7cy/indderived_W1_Anon_V7.0.0.dta?rlkey=y79rxp6euqe82pj3tlo7y8xmh&st=pamh08ef&dl=1',
    './data/indderived_w1.dta',
    mode = "wb"
)

w1_best <- read_dta('./data/indderived_w1.dta') # pull in data containing best gen variable in W1

w1_best_gender <- w1_best %>%
    filter(w1_quest_typ == 1) %>% # only want those who did adult questionnaire
    select(pid, w1_best_gen)

download.file(
    'https://www.dropbox.com/scl/fi/vkxbocpglujiix6fyd2m0/indderived_W4_Anon_V2.0.0.dta?rlkey=xwetrj2fmp75oozqh8bgo6xgz&st=0meb21px&dl=1',
    './data/indderived_w4.dta',
    mode = "wb"
)
w4_best <- read_dta('./data/indderived_w4.dta') # pull in data containing best gen variable in W4

w4_best_gender_empl <- w4_best %>%
    filter(w4_quest_typ == 1) %>% # only want those who did adult questionnaire
    select(pid, w4_best_gen, w4_empl_stat, w4_best_race)

# now want to merge the best gender and best employment status columns into the main data

clean_w1_best <- clean_w1 %>%
    left_join(w1_best_gender, by = "pid")

clean_w4_best <- clean_w4 %>%
    left_join(w4_best_gender_empl, by = "pid")

joined_w1_w4_best <- clean_w1_best %>%
    left_join(clean_w4_best, by = "pid")
```

```{r Common sense test with best gender}
test_best <- joined_w1_w4_best %>% 
    filter(!is.na(w1_best_gen) & !is.na(w4_best_gen)) %>% # removes entries where gender was missing in either w1 or in w4
    mutate(match = w1_best_gen == w4_best_gen) # returns TRUE/FALSE from logical test to see if the genders in W1 and W4 match for each observation

unique(test_best$match)
```

(WITH BEST GENDER VARIABLE) Now have 3 768 entries which were successfully joined. Once again the genders match. Now have only *380* missing observations for gender as opposed to *1 071*. W1 had no missing values reported for best gender, whereas *W4 had all the 380 missing values*. This might be explained by attrition post-W1.

## Create dummy variables coding missings for conscientiousness proxies

Now working towards running the OLS.

First, need to narrow down scope of proxies to 4 of the most important potential proxies. These are the 4 I've chosen:

-   J21: "How regularly do you exercise". Speaks to hard-working nature and self-discipline of individual. *(46/4 148 missing)*

-   H7: "Have you successfully completed any diplomas, certificates or degrees outside of school''. Can show resilience, ambition, and dutifulness. Would, however, have to control for income level. *(50/ 4148 missing)*

-   K5: "I felt hopeful about the future". Speaks to ambitiousness of person. *(47/ 4148 missing)*

### New potential questions:

-   H33: "Taking everything into account, do you intend to continue studying after matric, that is, after leaving school?". *(2046/ 4148 missing)* Too many missings and potential collinearity with question H7 makes it not usable.

-   J26: "Do you smoke cigarettes". *(31/ 4148 missing)* Measurement error and age group (14-23) makes this an uncertain measure.

-   M5: "Using a scale of 1 to 10 where 1 means “Very dissatisfied” and 10 means “Very satisfied”, how do you feel about your life as a whole right now?". *(498/ 4148 missing)*. Very weak correlate of conscientiousness I think.

- Maybe something to do with income such as investing or being in a stokvel. Would be tricky due to effect of level of income on this.

Now can create dummy variable coding missing values as 1 and valid entries as 0.

```{r Conscientiousness dummies}
# start by only selecting variables which we have narrowed study down to
clean_w1 <- clean_w1 %>% 
    select(pid, w1_a_hllfexer, w1_a_emohope, w1_a_edter, w1_a_edintter, w1_a_hllfsmk, w1_a_wbsat) %>% 
    # first want to create dummies for missing variables
    mutate(w1_a_hllfexer_miss = ifelse(w1_a_hllfexer %in% c(-9, -8, -5, -3) | is.na(w1_a_hllfexer), 1, 0)) %>% 
    mutate(w1_a_emohope_miss = ifelse(w1_a_emohope %in% c(-9, -8, -5, -3) | is.na(w1_a_emohope), 1, 0)) %>% 
    mutate(w1_a_edter_miss = ifelse(w1_a_edter %in% c(-9, -8, -5, -3) | is.na(w1_a_edter), 1, 0)) %>% 
    mutate(w1_a_edintter_miss = ifelse(w1_a_edintter %in% c(-9, -8, -5, -3) | is.na(w1_a_edintter), 1, 0)) %>% 
    mutate(w1_a_hllfsmk_miss = ifelse(w1_a_hllfsmk %in% c(-9, -8, -5, -3) | is.na(w1_a_hllfsmk), 1, 0)) %>%
    mutate(w1_a_wbsat_miss = ifelse(w1_a_wbsat %in% c(-9, -8, -5, -3) | is.na(w1_a_wbsat), 1, 0)) %>% 
    # now want to make sure that all these values get coded as missing in original columns
    mutate(w1_a_hllfexer = ifelse(w1_a_hllfexer %in% c(-9, -8, -5, -3), NA, w1_a_hllfexer)) %>% 
    mutate(w1_a_emohope = ifelse(w1_a_emohope %in% c(-9, -8, -5, -3), NA, w1_a_emohope)) %>% 
    mutate(w1_a_edter = ifelse(w1_a_edter %in% c(-9, -8, -5, -3), NA, w1_a_edter)) %>% 
    mutate(w1_a_edintter = ifelse(w1_a_edintter %in% c(-9, -8, -5, -3), NA, w1_a_edintter)) %>% 
    mutate(w1_a_hllfsmk = ifelse(w1_a_hllfsmk %in% c(-9, -8, -5, -3), NA, w1_a_hllfsmk)) %>% 
    mutate(w1_a_wbsat = ifelse(w1_a_wbsat %in% c(-9, -8, -5, -3), NA, w1_a_wbsat)) %>% 
    #also need to make sure that all variables are increasing in conscientiousness in same direction
    mutate(w1_a_edter_flipped = 3 - w1_a_edter)
```

## Create conscientiousness measure with PCA

-   Will need to use the `prcomp` function for this

-   Transforming data:

    -   Convert entries such that all go in same direction, i.e. that higher value indicates higher conscientiousness
    -   Some entries coded as -3 or 25 (these mean something). In this case, should I code as NA and then include them in my missing dummy.
    
```{r}
consc <- prcomp(na.omit(clean_w1[, c("w1_a_hllfexer",
                                     "w1_a_emohope",
                                     "w1_a_edter_flipped",
                                     "w1_a_hllfsmk")]), 
                center = T, 
                scale. = T)

# Create a vector of NAs with original length
pca_scores <- rep(NA, nrow(clean_w1))

# Get indices of complete cases used in PCA
complete_rows <- complete.cases(clean_w1[, c("w1_a_hllfexer",
                                             "w1_a_emohope",
                                             "w1_a_edter_flipped",
                                             "w1_a_hllfsmk")])

# Assign PCA scores only to complete cases
pca_scores[complete_rows] <- consc$x[, 1]

# Now assign back to original dataframe
clean_w1$w1_a_consc <- pca_scores



```

# Run preliminary OLS for proposal

This section runs preliminary OLS models for inclusion into my research proposal.

```{r OLS: W1 conscientiousness on W4 employment status}

joined_w1_w4_best <- joined_w1_w4_best %>%
    mutate(w4_empl_stat = ifelse(w4_empl_stat == -8, NA, w4_empl_stat)) %>%  # code those who refused to answer as missing (15 people)
    mutate(w4_employed = ifelse(w4_empl_stat == 3, 1, 0)) %>% 
    mutate(w4_best_gen = as.factor(w4_best_gen)) %>%  # coding gender as a categorical variable
    mutate(w4_best_race = as.factor(w4_best_race))

ols1 <- lm_robust(joined_w1_w4_best$w4_employed ~ clean_w1$w1_a_consc + joined_w1_w4_best$w4_best_gen + joined_w1_w4_best$w4_best_race)

print(ols1)
```


# Cleaning W4 education and Wages 

Hello world

