---
output: github_document
---
# Purpose

This README captures all my code which I had used in the creation of my thesis for 2025.

```{r Setup, include=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
pacman::p_load(tidyverse, foreign, haven, estimatr, broom, kableExtra, FactoMineR)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

# Running initial OLS on data

-   An OLS model was created to test if the variables are somewhat significant in the first place.
-   It helps to inform the general direction of the project and to establish any relationships from early on.

## Pull Data into R

```{r Pulling data into R}
# Adult questionnaires

    # W1
w1 <- read_dta('https://www.dropbox.com/scl/fi/3s45c3flubpx0ngbz4qjq/Adult_W1_Anon_V7.0.0.dta?rlkey=oa17mxoimb3gqdo7x7jh90elu&st=dpzn40ah&dl=1')

    # W4
w4 <- read_dta('https://www.dropbox.com/scl/fi/3ggfkjhj55z09i11e40ft/Adult_W4_Anon_V2.0.0.dta?rlkey=wjrlq33ggrohe5o8z3mpjsrda&st=ul2l4ay9&dl=1')

# Individual-derived data

    # W1
w1_best <- read_dta('https://www.dropbox.com/scl/fi/etluqr514bxyzigroy7cy/indderived_W1_Anon_V7.0.0.dta?rlkey=y79rxp6euqe82pj3tlo7y8xmh&st=pamh08ef&dl=1') 

    # W4
w4_best <- read_dta('https://www.dropbox.com/scl/fi/vkxbocpglujiix6fyd2m0/indderived_W4_Anon_V2.0.0.dta?rlkey=xwetrj2fmp75oozqh8bgo6xgz&st=0meb21px&dl=1')

# HH questionnaires

    # W1
w1_hh_roster <- read_dta("https://www.dropbox.com/scl/fi/xlziinwu2txydwv2drcz6/HouseholdRoster_W1_Anon_V7.0.0.dta?rlkey=wku2iqdxbuibeibt0vuqpi3iq&st=58xqewm5&dl=1")

w1_hh <- read_dta("https://www.dropbox.com/scl/fi/gnyzrt6ftftx7pn694501/HHQuestionnaire_W1_Anon_V7.0.0.dta?rlkey=xdbsbh790f0upvtztye0lmdzi&st=qw3bxzz4&dl=1")

w1_child <- read_dta("https://www.dropbox.com/scl/fi/q28aia37wu3ijhv49we5m/Child_W1_Anon_V7.0.0.dta?rlkey=c8t4niecef95cep5z4hues34p&st=j104v6e6&dl=1")
```

## Clean and wrangle data

-   Obtain W1 data and clean to include only variables which proxy for conscientiousness and other relevant variables

```{r Cleaning W1}
# first want to extract target group, those aged 14-23 in W1

targetpop <- get_target_pop(df = w1_best, 
                            lower_age = 14, 
                            upper_age = 23)

clean_w1 <- targetpop %>% 
    left_join(w1 %>% 
                 select(pid, w1_a_dob_m, w1_a_dob_y, w1_a_gen, w1_a_unemmn, w1_a_hllfexer, w1_a_emohope, w1_a_aspen, w1_a_edter, w1_a_hllfsmk,
                        w1_a_edintter, w1_a_wbsat, w1_a_hldifmon), by = "pid") # selecting our questions which proxied for measures of conscientiousness
    
```

-   Now want to clean W4 data to obtain W4 wages, employment, and schooling

```{r Cleaning W4}
clean_w4 <- w4 %>% 
    select(pid, w4_a_gen, w4_a_dob_y, w4_a_em1, w4_a_em1pay, w4_a_edschgrd) # collecting columns with info on employment status, wages and schooling level
```

Tips:

-   Include common sense test
-   Include dummy variable for missing values of conscientiousness, start with 4/5 most promising then add more if needed.
-   Use PCA/MCA for constructing measure. Do some research on this

## Joining W1 and W4

```{r Joining W1 and W4}
joined_w1_w4 <- clean_w1 %>% 
    left_join(clean_w4, by = "pid")
```

## Add a common sense test

This will help catch if there are errors in the data, e.g. does reported gender change from W1 to W4.

```{r Common sense test}
test1 <- joined_w1_w4 %>% 
    filter(!is.na(w1_a_gen) & !is.na(w4_a_gen)) %>% # removes entries where gender was missing in either w1 or in w4
    mutate(match = w1_a_gen == w4_a_gen) # returns TRUE/FALSE from logical test to see if the genders in W1 and W4 match for each observation

unique(test1$match) # checking if we have only TRUE values
```

Test shows that data is not flawed!! :)

(WITH OLD GENDER VARIABLE) 3 077 entries were successfully joined and had passed the common sense test. This is different from the `joined_w1_w4` sample size (4 148 observations) due to the exclusion of those entries with missing variables. 1 071 variables drop out because of a missing gender variable in either W1 or W4. Therefore must try with the `best_gen` variable as this provides a better measure for test.

## Trying common sense test with best gender

```{r Common sense test2}


w1_best_gender <- w1_best %>%
    filter(w1_quest_typ == 1) %>% # only want those who did adult questionnaire
    select(pid, w1_best_gen)



w4_best_gender_empl <- w4_best %>%
    filter(w4_quest_typ == 1) %>% # only want those who did adult questionnaire
    select(pid, w4_best_gen, w4_empl_stat, w4_best_race)

# now want to merge the best gender and best employment status columns into the main data

clean_w1_best <- clean_w1 %>%
    left_join(w1_best_gender, by = "pid")

clean_w4_best <- clean_w4 %>%
    left_join(w4_best_gender_empl, by = "pid")

joined_w1_w4_best <- clean_w1_best %>%
    left_join(clean_w4_best, by = "pid")
```

```{r Common sense test with best gender}
test_best <- joined_w1_w4_best %>% 
    filter(!is.na(w1_best_gen) & !is.na(w4_best_gen)) %>% # removes entries where gender was missing in either w1 or in w4
    mutate(match = w1_best_gen == w4_best_gen) # returns TRUE/FALSE from logical test to see if the genders in W1 and W4 match for each observation

unique(test_best$match)
```

(WITH BEST GENDER VARIABLE) Now have 3 768 entries which were successfully joined. Once again the genders match. Now have only *380* missing observations for gender as opposed to *1 071*. W1 had no missing values reported for best gender, whereas *W4 had all the 380 missing values*. This might be explained by attrition post-W1.

## Create dummy variables coding missings for conscientiousness proxies

Now working towards running the OLS.

First, need to narrow down scope of proxies to 4 of the most important potential proxies. These are the 4 I've chosen:

-   J21: "How regularly do you exercise". Speaks to hard-working nature and self-discipline of individual. *(46/4 148 missing)*

-   H7: "Have you successfully completed any diplomas, certificates or degrees outside of school''. Can show resilience, ambition, and dutifulness. Would, however, have to control for income level. *(50/ 4148 missing)*

-   K5: "I felt hopeful about the future". Speaks to ambitiousness of person. *(47/ 4148 missing)*

### New potential questions:

-   H33: "Taking everything into account, do you intend to continue studying after matric, that is, after leaving school?". *(2046/ 4148 missing)* Too many missings and potential collinearity with question H7 makes it not usable.

-   J26: "Do you smoke cigarettes". *(31/ 4148 missing)* Measurement error and age group (14-23) makes this an uncertain measure.

-   M5: "Using a scale of 1 to 10 where 1 means “Very dissatisfied” and 10 means “Very satisfied”, how do you feel about your life as a whole right now?". *(498/ 4148 missing)*. Very weak correlate of conscientiousness I think.

-   J24.7: "Indicate the level of difficulty you have with managing money (if you had to)"

- Maybe something to do with income such as investing or being in a stokvel. Would be tricky due to effect of level of income on this.

Now can create dummy variable coding missing values as 1 and valid entries as 0.

```{r Conscientiousness dummies}
# start by only selecting variables which we have narrowed study down to
clean_w1 <- clean_w1 %>% 
    select(pid, w1_a_hllfexer, w1_a_emohope, w1_a_edter, w1_a_edintter, w1_a_hllfsmk, w1_a_wbsat, w1_a_hldifmon) %>% 
    # create conscientiousness dummy
    consc_dummy(c("w1_a_hllfexer", "w1_a_emohope", "w1_a_edter", "w1_a_edintter", "w1_a_hllfsmk", "w1_a_wbsat", "w1_a_hldifmon")) %>% 
    # now want to make sure that all these values get coded as missing in original columns
    nids_miss(c("w1_a_hllfexer", "w1_a_emohope", "w1_a_edter", "w1_a_edintter", "w1_a_hllfsmk", "w1_a_wbsat", "w1_a_hldifmon"))%>% 
    #also need to make sure that all variables are increasing in conscientiousness in same direction
    mutate(w1_a_edter_flipped = 3 - w1_a_edter)
```

## Create conscientiousness measure with PCA

-   Will need to use the `prcomp` function for this

-   Transforming data:

    -   Convert entries such that all go in same direction, i.e. that higher value indicates higher conscientiousness
    -   Some entries coded as -3 or 25 (these mean something). In this case, should I code as NA and then include them in my missing dummy.
    
```{r Make conscientiousness and missing conscientiousness variables}
consc <- prcomp(na.omit(clean_w1[, c("w1_a_hllfexer",
                                     "w1_a_emohope",
                                     "w1_a_edter_flipped",
                                     "w1_a_hllfsmk")]), 
                center = T, 
                scale. = T)

# Create a vector of NAs with original length
pca_scores <- rep(NA, nrow(clean_w1))

# Get indices of complete cases used in PCA
complete_rows <- complete.cases(clean_w1[, c("w1_a_hllfexer",
                                             "w1_a_emohope",
                                             "w1_a_edter_flipped",
                                             "w1_a_hllfsmk")])

# Assign PCA scores only to complete cases
pca_scores[complete_rows] <- consc$x[, 1]

# Now assign back to original dataframe
clean_w1$w1_a_consc <- pca_scores
```

# Run preliminary OLS for proposal

This section runs preliminary OLS models for inclusion into my research proposal.

```{r OLS: W1 conscientiousness on W4 employment status}

joined_w1_w4_best <- joined_w1_w4_best %>%
    mutate(w4_empl_stat = ifelse(w4_empl_stat == -8, NA, w4_empl_stat)) %>%  # code those who refused to answer as missing (15 people)
    mutate(w4_employed = ifelse(w4_empl_stat == 3, 1, 0)) %>% 
    mutate(w4_best_gen = as.factor(w4_best_gen)) %>%  # coding gender as a categorical variable
    mutate(w4_best_race = as.factor(w4_best_race)) # coding race as a categorical variable

joined_w1_w4_best$w1_a_consc <- pca_scores



```


# Continue work on OLS

The following chunks are my post-proposal code and serve to continue the work to getting to an OLS model with additional controls.

## Create wealth measure (asset index) with MCA

Use MCA to construct a measure of wealth of individuals. This measure can then be used as a control in the model.

```{r Create wealth MCA}
# pull variables into df which give us info on asset ownership
# clean_w1 <- clean_w1 %>%  
#     left_join(select(w1, pid, w1_a_ownrad, w1_a_ownvehpri, w1_a_ownmot, w1_a_owncom, w1_a_owncel), by = "pid") %>% 
#     nids_miss(c("w1_a_ownrad", "w1_a_ownvehpri", "w1_a_ownmot", "w1_a_owncom", "w1_a_owncel"))
# 
# # testing to see if my mutate command worked to code any `2`, i.e. `No`, responses as 0 and leaving 1 = yes     
# test_mut <- clean_w1 %>% 
#     mutate(across(c("w1_a_ownrad", "w1_a_ownvehpri", "w1_a_ownmot", "w1_a_owncom", "w1_a_owncel"), ~ ifelse(.x == 2, 0, .x)))

######Stopped here ##########
# Select the asset ownership columns
# asset_vars <- df %>%
#   select(car, radio, cellphone, computer, fridge, tv)
# 
# # Convert asset vars to factors (MCA requires categorical variables)
# asset_factors <- asset_vars %>%
#   mutate(across(everything(), as.factor))
# 
# # Run MCA
# mca_result <- MCA(asset_factors, graph = FALSE)
# 
# # Extract the coordinates of individuals on the first dimension
# asset_index <- mca_result$ind$coord[, 1]
# 
# # Add the asset index as a new column to your original dataframe
# df <- df %>%
#   mutate(asset_index = asset_index)


wealth <- create_wealth_index(w1_best)

wealth_age <- wealth %>% left_join(w1_best %>% select(pid, w1_best_age_yrs), by = "pid") %>% mutate(l_total_assets_pc = log(total_assets_pc))

joined_w1_w4_best <- joined_w1_w4_best %>% 
     left_join(wealth_age %>% select(pid, l_total_assets_pc), by = "pid")

ols1 <- lm_robust(w4_employed ~ w1_a_consc 
                  + w4_best_gen + w4_best_race + l_total_assets_pc,
                   data = joined_w1_w4_best)
 summary(ols1) 


# Create a tibble with estimates, p-values, and significance stars

table_1 <- tidy(ols1) %>%
  filter(!is.na(estimate)) %>%  # remove dropped coefficients
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE            ~ ""
    ),
    estimate = round(estimate, 3),
    estimate = paste0(estimate, stars),
    std.error = round(std.error, 3),
    p.value = signif(p.value, 3),
    std.error = paste0("(", std.error, ")")
  ) %>%
  select(term, estimate, std.error, p.value) 

write_rds(table_1, "./output/tab1.rds")


 
 # create total income variable for W4
 w4_best <- w4_best %>% mutate(total_inc = rowSums(.[, 17:42], na.rm = T))
 
joined_w1_w4_best <- joined_w1_w4_best %>% 
    left_join(w4_best %>% select(pid, total_inc), by = "pid")

joined_w1_w4_best <- joined_w1_w4_best %>% 
    mutate(l_total_inc = log(total_inc)) %>% 
    mutate(l_total_inc = ifelse(l_total_inc == "-Inf", NA, l_total_inc)) %>% 
    mutate(l_total_assets_pc = ifelse(l_total_assets_pc == "-Inf", NA, l_total_assets_pc)) %>% 
    mutate(conscientiousness_shifted = w1_a_consc + if_else(min(w1_a_consc, na.rm = TRUE) < 0,
                                                                  abs(min(w1_a_consc, na.rm = TRUE)),
                                                                  0))


 
 ols2 <- lm(l_total_inc ~ w1_a_consc + w4_best_gen + w4_best_race + l_total_assets_pc,
                  data = joined_w1_w4_best)

 table_2 <- tidy(ols2) %>%
  filter(!is.na(estimate)) %>%  # remove dropped coefficients
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE            ~ ""
    ),
    estimate = round(estimate, 3),
    estimate = paste0(estimate, stars),
    std.error = round(std.error, 3),
    p.value = signif(p.value, 3),
    std.error = paste0("(", std.error, ")")
  ) %>%
  select(term, estimate, std.error, p.value) 
 
 write_rds(table_2, "./output/tab2.rds")
 
summary(ols2)
```


# Revised OLS

Want to include some revisions to the original models run to provide potentially better estimates.

## Include age and educ in df

```{r}
# rename df to ols_df for ease of recalling

# adding age information
ols_df <- joined_w1_w4_best %>% 
    left_join(w1_best %>% select(pid, w1_best_age_yrs),
              by = "pid") %>% # add age column
    mutate(age_sq = w1_best_age_yrs^2) %>%  # create age squared column
    rename("w1_age" = "w1_best_age_yrs") %>% 
    mutate(w1_age_groups = case_when(w1_age >=14 & w1_age <= 15 ~ "14-15",
                                  w1_age >= 16 & w1_age <=19 ~ "16-19",
                                  w1_age >= 20 ~ "20-23")) # create age groupings 

# adding education information
ols_df <- ols_df %>% 
    left_join(w1_best %>% select(pid, w1_best_edu), 
              by = "pid") %>% 
    nids_miss("w1_best_edu") %>% # code missings as NA
    mutate(w1_best_edu = ifelse(w1_best_edu == 24,
                                NA,
                                w1_best_edu)) %>%  # code those who reported "Other" for level of educ as missing
    mutate(w1_best_edu = ifelse(w1_best_edu == 25,
                                0,
                                w1_best_edu + 1))

```


## Factor Analysis of Mixed Data

First create df of only asset values and binary ownership columns

```{r}

```

