---
title: "Meeting_10_07_2026"
output: github_document
---

```{r Setup, include=FALSE}
library(kableExtra)
library(tidyverse)
```



# Recap

- Trying to apply same method of optimisation from reduced sample (only black males) to full sample, and with controls
- Had 2 hypotheses:
    1. Nothing fixed with optimisation
    2. Local optimiser found the global min

- In response to this, I removed init_par and found that the parameters were did not minimise loss value as well as local optim function did
- Below I set out the results of the local optim function which were then parsed into global mbo functions:

# Results



```{r echo=FALSE}
model_with_controls <- readRDS(file = '/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/mbo_control.rds')

model_simple <- readRDS(file = '/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/mbo_simple.rds')

model_simple <- model_simple %>% 
    mutate(Parameter_meaning = case_when(
        Parameter == "beta_0_educ"   ~ "Intercept term of educ equation",
        Parameter == "alpha_educ"    ~ "Consc on educ",        
        Parameter == "beta_0_educ"   ~ "Intercept term of educ equation",
        Parameter == "beta_0_income" ~ "Intercept term of income equation",
        Parameter == "gamma_income"  ~ "Educ on income",
        Parameter == "alpha_income"  ~ "Consc on income",
        Parameter == "u_consc"       ~ "Std dev of consc error term",
        Parameter == "u_educ"        ~ "Std dev of educ equation error term",
        Parameter == "u_income"      ~ "Std dev of income equation error term",
        Parameter == "delta_income_white"   ~ "White on income",
        Parameter == "delta_income_coloured"   ~ "Coloured on income",
        Parameter == "delta_income_asian"   ~ "Asian on income",
        Parameter == "delta_income_female"   ~ "Female on income",
        Parameter == "delta_income_age"   ~ "Age on income",
        Parameter == "delta_income_age_squared" ~ "Intercept term of educ equation",
        Parameter == "delta_income_educ_squared" ~ "Intercept term of educ equation",
        Parameter == "delta_income_wealth" ~ "Wealth on income",
        TRUE ~ NA_character_
    ))

model_with_controls %>% 
    mutate(Parameter_meaning = case_when(
        Parameter == "beta_0_educ"   ~ "Intercept term of educ equation",
        Parameter == "alpha_educ"    ~ "Consc on educ",        
        Parameter == "beta_0_educ"   ~ "Intercept term of educ equation",
        Parameter == "beta_0_income" ~ "Intercept term of income equation",
        Parameter == "gamma_income"  ~ "Educ on income",
        Parameter == "alpha_income"  ~ "Consc on income",
        Parameter == "u_consc"       ~ "Std dev of consc error term",
        Parameter == "u_educ"        ~ "Std dev of educ equation error term",
        Parameter == "u_income"      ~ "Std dev of income equation error term",
        Parameter == "delta_income_white"   ~ "White on income",
        Parameter == "delta_income_coloured"   ~ "Coloured on income",
        Parameter == "delta_income_asian"   ~ "Asian on income",
        Parameter == "delta_income_female"   ~ "Female on income",
        Parameter == "delta_income_age"   ~ "Age on income",
        Parameter == "delta_income_age_squared" ~ "Intercept term of educ equation",
        Parameter == "delta_income_educ_squared" ~ "Intercept term of educ equation",
        Parameter == "delta_income_wealth" ~ "Wealth on income",
        TRUE ~ NA_character_
    )) %>% 
    left_join(model_simple, 
              by = "Parameter") %>% 
    select(Parameter, Value.y, Value.x) %>% 
    mutate(Value.y = ifelse(is.na(Value.y), "", Value.y)) %>% 
    rename("Simple" = "Value.y",
           "Controls" = "Value.x") %>% 
    kable(format = "html",
          caption = "Simulation result on full model with controls")
```

