---
title: "Meeting_27_03_2026"
output: github_document
---
```{r include=FALSE}
library(tidyverse)
```


# Progress

1. Fixed income specification in structural simulation function
    - Removed log from RHS

2. Increased sample size

3. Added controls to structural model

4. Applied weights in calculation of loss function
    - Weighted moments (mean and standard deviation) of variables by inverse of variance
    - Weighted coefficients by inverse of squared standard error

5. Used `mlrMBO` package to find global optima (scripts for simple and multivariate models found [here]() and [here]())

    - Simple results
    
```{r, echo=FALSE, results='markup'}
mbo_results_simple <- readRDS('/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/mbo_simple.rds')

mbo_results_simple
```

    - With controls

```{r, echo=FALSE, results='markup'}
mbo_results_controls <- readRDS('/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/mbo_control.rds')

mbo_results_controls
```



5. Estimated relationships using MLE
```{r, echo=FALSE, results='markup'}

ml_tab <- readRDS(file = '/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/ml_tab.rds')

# convert to dataframe
ml_results_df <- ml_tab %>%
  rownames_to_column(var = "parameter") %>%
  select(parameter, Estimate, `Std. error`, `t value`, `Pr(> t)`) %>%
  mutate(
    stars = case_when(
      `Pr(> t)` < 0.001 ~ "***",
      `Pr(> t)` < 0.01  ~ "**",
      `Pr(> t)` < 0.05  ~ "*",
      `Pr(> t)` < 0.1   ~ ".",
      TRUE              ~ ""
    )
  ) %>%
  mutate(
    estimates = paste0(round(Estimate, 3), stars)
  ) %>% 
    select(parameter, estimates) %>% 
    mutate(parameter = case_when(
        parameter == "alpha_income" ~ "alpha_income(consc)",
        parameter == "gamma_income" ~ "gamma_income(educ)",
        TRUE ~ parameter
    ))

print(ml_results_df)
```

