---
title: "Meeting_21_05_2026"
output: github_document
---

# Recap

- Structural simulation:
    - Getting unrealistic results in both simple and multiple-controls specifications of structural modelling
- Max Likelihood:
    - Getting NA for standard errors of intercept term (fixed)


# Progress

- Updated code to run simulation on sample of only black males
- Incorporated random design and design matrix into estimation
- Get result that optimal parameter values are the initial values specified??
    - Potentially stems from misspecified `mbo` code:
    
Code currently says:

```{r}
# library(mlrMBO)
# library(ParamHelpers)
# library(parallelMap)
# 
# 
# 
# # configure Bayesian optimisation
# mbo_ctrl <- makeMBOControl()
# mbo_ctrl <- setMBOControlInfill(mbo_ctrl, crit = crit.ei)      # expected improvement
# mbo_ctrl <- setMBOControlTermination(mbo_ctrl, max.evals = 500L) # iterations
# 
# # generate initial values and random design
# 
# init_par <- data.frame(
#     beta_0_educ               = 10,
#     alpha_educ                = 0.5,
#     beta_0_income             = 5,
#     gamma_income              = 0.3,
#     alpha_income              = 0.2,
#     u_consc                   = 0,
#     u_educ                    = 0,
#     u_income                  = 0
# )
# 
# random_design <- generateRandomDesign(n = 20, par.set = par_set_simple)
# 
# design_mat <- rbind(init_par, random_design)
# 
# 
# ctrl <- mlr::makeTuneControlMBO(mbo.control = mbo_ctrl, mbo.design = design_mat)
# 
# # run in parallel
# parallelStartSocket(cpus = parallel::detectCores() - 1)
# 
# parallelExport(
#     "sim_function_reg",
#     "target_values_simple_black_males",
#     "weights_vec"
# )
# 
# # also load required packages on each worker
# parallelLibrary("tidyverse")
# parallelLibrary("mlrMBO")
# 
# set.seed(427292, "L'Ecuyer")
# result <- mbo(
#     fun     = obj_fun,
#     design  = design_mat,
#     control = mbo_ctrl,
#     show.info = TRUE
# )
# 
# 
# parallelStop()
# 
# # extract results
# result$x        # optimal parameter values
# result$y        # final loss value

```

Code from tutorial 12 looks like this:

```{r}
# Setup: defining the learning task in mlr
# regr_task <- mlr::makeRegrTask(
#   id = "NIDS income regression with elasticnet",
#   data = df_train_pre_proc, 
#   target = "log_income_pc"
#   )
# 
# # Setup: Baysesian optimisation control (mlrMBO)
# measure <- rmse
# mbo_ctrl <- mlrMBO::makeMBOControl()
# mbo_ctrl <- mlrMBO::setMBOControlInfill(mbo_ctrl, crit = crit.ei)
# mbo_ctrl <- mlrMBO::setMBOControlTermination(mbo_ctrl, max.evals = 50L)
# 
# # Setup: learner and hyperparameter space
# regr_lrn <- mlr::makeLearner("regr.glmnet")
# glmnet_num_ps = ParamHelpers::makeParamSet(
#   ParamHelpers::makeNumericParam("alpha", lower = 0, upper = 1),
#   ParamHelpers::makeNumericParam("lambda", lower = 0, upper = 0.1)
# )
# 
# # Tuning execution
# design_mat <- ParamHelpers::generateRandomDesign(n = 8, par.set = glmnet_num_ps) 
# ctrl <- mlr::makeTuneControlMBO(mbo.control = mbo_ctrl, mbo.design = design_mat)
# 
# set.seed(427292, "L'Ecuyer")
# # parallelStartSocket(min(6, detectCores() - 1))
# tune_pars_glmnet <- mlr::tuneParams(
#   learner = regr_lrn, 
#   task = regr_task, 
#   resampling = mlr::makeResampleDesc("CV", iters = 10L),
#   measures = rmse, 
#   par.set = glmnet_num_ps, 
#   control = ctrl, 
#   show.info = TRUE
#   )
# parallelStop()

```

