---
title: "Meeting_05_06_2026"
output: github_document
---

```{r include=FALSE}
library(tidyverse)
```


# Progress

- Investigated why I was getting fed back initial values as results from MBO output:   
    - Seemed that initial values specified were already values which minimised loss  

Print optimal path results here for 10 best iterations, in terms of smallest loss values
Using first set of initial values (Where `y` here corresponds to loss value):

```{r}
init_par <- data.frame(
    beta_0_educ   = 10,  # educ intercept
    alpha_educ    = 0.5, # consc on educ
    beta_0_income = 5,   # income intercept
    gamma_income  = 0.3, # educ on income
    alpha_income  = 0.2, # consc on income
    u_consc       = 0,   # sd of consc 
    u_educ        = 0,   # sd of educ error
    u_income      = 0    # sd of income error
)
```

```{r echo=FALSE}

opt_path1 <- readRDS("/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/optpath1.rds")

opt_path_out <- head(opt_path1[order(opt_path1$y), ], 10) 

opt_path_out %>% 
    mutate(across(1:8, ~ round(.x, 3))) %>% 
    select(1:9)
```

Using second set of initial values:
```{r}
init_par <- data.frame(
    beta_0_educ               = 1,
    alpha_educ                = 0.2,
    beta_0_income             = 8,
    gamma_income              = 0.4,
    alpha_income              = 0.5,
    u_consc                   = 1,
    u_educ                    = 1,
    u_income                  = 1
)
```

```{r echo=FALSE}
opt_path2 <- readRDS("/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/optpath2.rds")

opt_path_out2 <- head(opt_path2[order(opt_path2$y), ], 10) 

opt_path_out2 %>% 
    mutate(across(1:8, ~ round(.x, 3))) %>% 
    select(1:9)
```


Then ran an optimisation function to try and generate values which would minimise `y` then fed these into `mbo` as initial values:
```{r}
# Minimisation code I ran to get new set of initial values:

# optim_result <- optim(
#     par     = c(10, 0.5, 5, 0.3, 0.2, 0, 0, 0),  # rough starting guess
#     fn      = function(par) {
#         sim_function_reg(
#             par          = par,
#             n            = 10000,
#             target_coefs = target_values_simple_black_males,
#             weights      = weights_vec
#         )
#     },
#     method  = "Nelder-Mead",
#     control = list(maxit = 1000, reltol = 1e-8)
# )

# above code came up with following set of parameter values to be parsed as initial values into `mbo` function:

init_par <- data.frame(
    beta_0_educ               = 11.50333,
    alpha_educ                = -0.02113,
    beta_0_income             = 6.361638,
    gamma_income              = 0.084179,
    alpha_income              = 0.04725592,
    u_consc                   = -0.4193541,
    u_educ                    = 1.310216,
    u_income                  = 0.03502108
)
```



```{r echo=FALSE}
opt_path3 <- readRDS("/Users/matthewthompson/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/optpath3.rds")

opt_path_out3 <- head(opt_path3[order(opt_path3$y), ], 10) 

opt_path_out3 %>% 
    mutate(across(1:8, ~ round(.x, 3))) %>% 
    select(1:9)
```

