---
title: "Meeting 1 August 2025"
author: "Matthew"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(kableExtra, tidyverse)
```

# Key points

- Constructed wealth index function
    - Using the parent id variables, I matched the monetary asset values of the parents to their children from our target group.
    - Code for this function found [here](https://github.com/mattthomo/Thesis_data_work/blob/main/code/create_wealth_index.R)
- Got OLS estimates for a model with log income (model 1) and employment status (model 2) both from wave 4 as dependent variables.
- 

# Results

## Model 1

```{r Model 1, results='asis'}
tab1 <- read_rds("~/Documents/Stellenbosch University/Masters/Research Assignment/Data_work/output/tab1.rds")

tab1 %>% kable(format = "html",
               booktabs = T,
               caption = "W1 Conscientiousness on W4 (log) income") %>% 
    row_spec(3, bold = T)

```

## Model 2




# Questions

- As results not significant and/or show relationship in direction opposite to what we had expected, next step?
    - Had lots of observations dropped due to missing entries
    - Still go ahead with more complexed estimation techniques?
    - Go back and change specification of conscientiousness variable, or find source of missingness
    - Simulations? (how to I'm not sure but sounded promising after yesterday's lecture)


